---
title: "R Coursework"
author: "Michelle"
date: "11/16/2021"
output: html_document
---

```{r}
#Load library
library(dplyr)
library(ggplot2)
library(tidyverse)
library(hrbrthemes)
library(zoo)
library(ggrepel)
library(skimr)
library(mlr3)
library(mlr3learners)
library(glmnet)
library(reshape)
library(ggfortify)
library(mlr3viz)
```

```{r}
#Load the data from the csv files

airports <- read.csv("airports.csv", header = TRUE)
carriers <- read.csv("carriers.csv", header = TRUE)
planes <- read.csv("plane-data.csv", header = TRUE)

flight2004 <- read.csv("2004.csv")
flight2005 <- read.csv("2005.csv")
```

```{r}
#Row binding the data from year 2004 and 2005
flights <- rbind(flight2004, flight2005)
```

```{r}
#Clean Data
flights <- flights %>%
  filter(Cancelled == 0 & Diverted == 0) %>%
  distinct()
```

Question 1
**When is the best time of day, day of the week, and time of year to fly to minimize delays?**
```{r} 
#Group CRSDepTime in 2 hours interval
flights <- flights %>%
  mutate(interval_deptime = case_when(
    CRSDepTime >= 000 & CRSDepTime < 200 ~ "00.00-01.59",
    CRSDepTime >= 200 & CRSDepTime < 400 ~ "02.00-03.59",
    CRSDepTime >= 400 & CRSDepTime < 600 ~ "04.00-05.59",
    CRSDepTime >= 600 & CRSDepTime < 800 ~ "06.00-07.59",
    CRSDepTime >= 800 & CRSDepTime < 1000 ~ "08.00-09.59",
    CRSDepTime >= 1000 & CRSDepTime < 1200 ~ "10.00-11.59",
    CRSDepTime >= 1200 & CRSDepTime < 1400 ~ "12.00-13.59",
    CRSDepTime >= 1400 & CRSDepTime < 1600 ~ "14.00-15.59", 
    CRSDepTime >= 1600 & CRSDepTime < 1800 ~ "16.00-17.59",
    CRSDepTime >= 1800 & CRSDepTime < 2000 ~ "18.00-19.59",
    CRSDepTime >= 2000 & CRSDepTime < 2200 ~ "20.00-21.59",
    CRSDepTime >= 2200 & CRSDepTime < 2400 ~ "22.00-23.59")
  )
flights
```

```{r}
best_time_of_day <- flights %>%
  filter(ArrDelay > 0) %>%
  group_by(interval_deptime) %>%
  summarise(avg_delay = mean(ArrDelay)) %>%
  arrange(interval_deptime)
best_time_of_day
```

```{r}
ggplot(best_time_of_day, aes(x = interval_deptime, y = avg_delay, 
                             label=sprintf("%0.2f", round(avg_delay, digits = 2)))) + 
  geom_bar(stat = "identity", color = "black", fill = "lightblue", width = 0.8) +
  labs(title = "Best Time Of Day", x = "Departure Time", y = "Average Delay (min)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(size = 3, hjust = 1.1, vjust = 0.5) +
  coord_flip()
```

```{r}
best_day_of_week <- flights %>%
  filter(ArrDelay > 0) %>%
  group_by(DayOfWeek) %>%
  summarise(avg_delay = mean(ArrDelay)) %>%
  arrange(DayOfWeek)
best_day_of_week
```

```{r}
best_day_of_week$DayOfWeek <- factor(best_day_of_week$DayOfWeek,levels = c("1", "2", "3", "4","5","6","7"))
ggplot(best_day_of_week, aes(x = DayOfWeek, y = avg_delay, 
                             label=sprintf("%0.2f", round(avg_delay, digits = 2)))) +
  geom_bar(stat = "identity", color = "black", fill = "lightblue", width = 0.7) +
  scale_x_discrete(labels=c("1" = "Monday", "2" = "Tuesday", "3" = "Wednesday", "4" = "Thursday", "5"="Friday", "6"="Saturday","7"="Sunday")) +
  labs(title = "Best Day Of Week", x= "Day", y = "Average Delay (min)") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(size = 4, hjust = 0.5, vjust = 1.5)
```

```{r}
best_time_of_year <- flights %>%
  filter(ArrDelay > 0) %>%
  group_by(Month) %>%
    summarise(avg_delay = mean(ArrDelay)) %>%
  arrange(Month)
best_time_of_year
```

```{r}
best_time_of_year$Month <- factor(best_time_of_year$Month,levels = c("1", "2", "3", "4","5","6","7", "8", "9", "10", "11", "12"))
ggplot(best_time_of_year, aes(x = Month, y = avg_delay, label=sprintf("%0.2f", round(avg_delay, digits = 2)))) + 
  geom_bar(stat = "identity", color = "black", fill = "lightblue", width = 0.8) +
  scale_x_discrete(labels=c("1" = "Jan", "2" = "Feb", "3" = "Mar", "4" = "Apr", "5"="May", "6"="June","7"="July", "8"="Aug", "9" = "Sep", "10"= "Oct", "11"="Nov", "12"= "Dec")) +
  labs(title = "Best Time Of Year", x = 'Month', y = 'Average Delay (min)') +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(size = 3, hjust = 0.5, vjust = 1.5)
```

Question 2
**Do older planes suffer more delays?**  
```{r}
#Inner join flights and planes by tailnum
planes_age <- flights %>%
  inner_join(planes, by = c("TailNum" = "tailnum")) %>%
  filter(ArrDelay > 0) %>%
  group_by(year) %>%
  summarise(avg_delay = mean(ArrDelay), total_flights = n()) %>%
  arrange(year)
planes_age
```

```{r}
planes_age <- na.omit(planes_age) %>%
  filter(year != "None" & year != "" & year != "0000" & year <= 2005)
planes_age
```

```{r}
ggplot(planes_age, aes(x= as.numeric(year), y=avg_delay)) +
  geom_point(shape=21, color="black", fill="#69b3a2", size=2) +
  geom_smooth(method = lm, se = FALSE, color = "darkgrey") +
  labs(title = "Average Arrival Delay Based on Aircraft's Age", 
       x= "Year Manufactured", y = "Average Delay (min)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_x_continuous(breaks=c(1955,1965,1975,1985,1995,2005))
```

```{r}
older_planes <- planes_age %>%
  filter(year >= 1956 & year <= 1980)

#Calculate Total Flights for Older Planes
sum(older_planes$total_flights)
```

```{r}
younger_planes <- planes_age %>%
  filter(year >= 1981 & year <= 2005)

#Calculate Total Flights for Younger Planes
sum(younger_planes$total_flights)
```

**Since the sample size for older planes is relatively low compared to younger planes, the estimation would be less accurate. Hence we plot the younger planes with a larger sample size to provide more accurate results.**

```{r}
ggplot(younger_planes, aes(x=as.numeric(year), y=avg_delay)) +
  geom_point(shape=21, color="black", fill="#69b3a2", size=2) +
  geom_smooth(method = lm, se = FALSE, color = "darkgrey") +
  labs(title = "Average Arrival Delay Of Younger Planes", x= "Year Manufactured", y = "Average Delay (min)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_x_continuous(breaks=c(1981,1985,1990,1995,2000,2005))
```
Question 3
**How does the number of people flying between different locations change over time?**
```{r}
#Combine Origin and Destination into Route
flights <- flights %>%
  mutate(Route = paste(Origin, Dest, sep = "-"))
flights
```

```{r}
#Count route in 2004
route_2004 <- flights %>%
  filter(grepl("2004", Year)) %>%
  count(Route)
  names(route_2004) <- c("Route", "2004")
route_2004
```
```{r}
#Count route in 2005
route_2005 <- flights %>%
  filter(grepl("2005", Year)) %>%
  count(Route)
  names(route_2005) <- c("Route", "2005")
route_2005
```
```{r}
#Inner join route_2004 and route_2005 so we can compare total flights between 2004 and 2005
route_merge <- route_2004 %>%
  inner_join(route_2005, by = "Route")
route_merge
```
```{r}
#Difference of Total Flight between 2004 and 2005 (2005-2005)
route_merge$Difference <- (route_merge$"2005" - route_merge$"2004")
route_merge
```

```{r}
#Top 5 of Route's Total Flights' Difference
top_route_merge <- route_merge %>%
  arrange(-Difference)
top_route_merge <- head(top_route_merge, 5)
top_route_merge
```
```{r}
#To make the visualization easier, we melt the data frame
top_route_merge_new <- top_route_merge %>%
  select("Route", "2004", "2005")
top_route_melt <- melt(top_route_merge_new)
names(top_route_melt) <- c("Route", "Year", "Total_Flights")
top_route_melt
```

```{r}
#Visualize Top 5 of Route's Total Flights' Difference
ggplot(top_route_melt, aes(Route, Total_Flights, fill= Year)) + 
    geom_bar(position="dodge", stat="identity") +
    labs(title = "Top 5 of Route's Total Flights' Difference Over Time", 
         x = "Route", y = "Total Flights") +
    theme(plot.title = element_text(hjust = 0.5)) +
    geom_text(aes(label = round(Total_Flights)), position = position_dodge(0.9),
              color="black",vjust = 1.5,hjust = 0.5, size = 2.5) +
    scale_y_continuous(labels = scales::comma)
```

```{r}
#Bottom 5 of Route's Total Flights' Difference
bottom_route_merge <- route_merge %>%
  arrange(Difference)
bottom_route_merge <- head(bottom_route_merge, 5)
bottom_route_merge
```

```{r}
#To make the visualization easier, we melt the data frame
bottom_route_merge_new <- bottom_route_merge %>%
  select("Route", "2004", "2005")
bottom_route_melt <- melt(bottom_route_merge_new)
names(bottom_route_melt) <- c("Route", "Year", "Total_Flights")
bottom_route_melt
```

```{r}
#Visualize Bottom 5 of Route's Total Flights' Difference
ggplot(bottom_route_melt, aes(Route, Total_Flights, fill= Year)) + 
    geom_bar(position="dodge", stat="identity") +
    labs(title = "Bottom 5 of Route's Total Flights' Difference Over Time", 
         x = "Route", y = "Total Flights") +
    theme(plot.title = element_text(hjust = 0.5)) +
    geom_text(aes(label = round(Total_Flights)), position = position_dodge(0.9),
              color="black",vjust = 1.5,hjust = 0.5, size = 2.5) +
    scale_y_continuous(labels = scales::comma)
```

Question 4
**Can you detect cascading failures as delays in one airport create delays in others?**
```{r}
cascadingfailures <- flights %>%
  inner_join(airports, by = c("Origin" = "iata")) %>%
  select(Year, Month, DayofMonth, DayOfWeek, DepTime, CRSDepTime, ArrTime, CRSArrTime, TailNum, ArrDelay, 
           DepDelay, Origin, Dest, airport) 
cascadingfailures
```

```{r}
delay <- cascadingfailures %>%
  filter(DepDelay > 100, ArrDelay > 100, Month == "12", DayofMonth == "18", TailNum == "N509US") %>%
  arrange(CRSDepTime)
delay
```

```{r}
ordered_airport  = factor(delay$airport, levels=c("Orlando International", "Detroit Metropolitan-Wayne County", "LaGuardia"))

ggplot()+
  geom_line(data=delay,aes(y=CRSArrTime,x= ordered_airport,colour="CRSArrTime"),group=1 )+
  geom_point(data=delay,aes(y=CRSArrTime,x= ordered_airport,colour="CRSArrTime")) +
  geom_line(data=delay,aes(y=ArrTime,x= ordered_airport,colour="ArrTime"),group=1) +
  geom_point(data=delay,aes(y=ArrTime,x= ordered_airport,colour="ArrTime")) +
  labs(title = "Delays in One Airport Create Delays in Others", x = "Airports", y = "Time")+
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(name = "Lines", values = c("CRSArrTime" = "darkblue", "ArrTime" = "red"))
```
Question 5
**Use the available variables to construct a model that predicts delays.**
```{r}
skim(flights)
```

```{r}
#Select variables to use for predicting delays
flight_model <- flights %>%
  select(Year, Month, DayofMonth, CRSDepTime, CRSArrTime, DepDelay, ArrDelay, FlightNum, Distance)
set.seed(1)
sample_flight <- sample_n(flight_model, 100000)
```

```{r}
#Set up a task
task <- TaskRegr$new("sample_flight", backend = sample_flight, target = "ArrDelay")
measure <- msr("regr.mse")
```

```{r}
#Split train and test set
set.seed(100)
set_train <- sample(task$nrow, 0.7 * task$nrow)
set_test <- setdiff(seq_len(task$nrow), set_train)
```

```{r}
#Using linear regression
learner_lm <- lrn("regr.lm")
learner_lm$train(task, row_ids = set_train)
learner_lm$predict(task, row_ids = set_test)$score()
```
```{r}
#Using Ridge Regression
learner_ridge <- lrn("regr.glmnet")
learner_ridge$param_set$values <- list(alpha = 0, lambda = 0.04)
learner_ridge$train(task, row_ids = set_train)
learner_ridge$predict(task, row_ids = set_test)$score()
```

```{r}
#Using Lasso Regression
learner_lasso <- lrn("regr.glmnet")
learner_lasso$param_set$values <- list(alpha = 1, lambda = 0.04)
learner_lasso$train(task, row_ids = set_train)
learner_lasso$predict(task, row_ids = set_test)$score()
```

```{r}
#Using Random Forest
learner_rf <- lrn("regr.ranger")
learner_rf$param_set$values <- list(min.node.size = 4)
learner_rf$train(task, row_ids = set_train)
learner_rf$predict(task, row_ids = set_test)$score()
```

```{r}
linear <- learner_lm$predict(task, row_ids = set_test)$score()
ridge <- learner_ridge$predict(task, row_ids = set_test)$score()
lasso <- learner_lasso$predict(task, row_ids = set_test)$score()
rf <- learner_rf$predict(task, row_ids = set_test)$score()
```


```{r}
#Comparing MSE between Linear Regression, Ridge Regression, Lasso Regression, and Random Forest
method <- data.frame (Method  = c("Linear Regression", "Ridge Regression", "Lasso Regression", "Random Forest"),
                      MSE = c(linear, ridge, lasso, rf)
                      )
method
```

```{r}
ggplot(method, aes(x = Method, y = MSE, label=sprintf("%0.2f", round(MSE, digits = 2)))) + 
  geom_bar(stat = "identity", color = "black", fill = "orange", width = 0.5) +
  labs(title = "Models", x = "Methods", y = "Mean Squared Error (MSE)") +
  theme(plot.title = element_text(hjust = 0.55)) +
  geom_text(size = 5, hjust = 0.5, vjust = 1.5)
```

